---
title: "Explore IBMR Data"
format: html
editor: visual
---


Explore X2 and Zooplankton inputs, lambda and minimum abundance outputs from SFHA 2025 IBMR effort

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, include = FALSE}
## Read in packages
rm(list=ls(all=TRUE))
library(tidyverse)
library(here)
library(plotly)
```


```{r, include = FALSE}
## Set paths
input_data_path <- here::here("data/data_processed")
output_path <- here::here("output/model_outputs/")
plot_path <- here::here("output/figures/")
```

```{r, include = FALSE}
## Read in data for all
x2a <- read_csv(here(input_data_path, "IBMR_X2_SF2025_input.csv")) %>%
  mutate(hydro = "AdjHist")
zoopa <- read_csv(here(input_data_path, "zoop_scalar_output_SF2025_2025-03-13.csv"))%>%
  mutate(hydro = "AdjHist")
x2b <- read_csv(here(input_data_path, "IBMR_X2_SF2022MED_input.csv"))%>%
  mutate(hydro = "2022MED")
zoopb <- read_csv(here(input_data_path, "zoop_scalar_output_SF2022MED_2025-03-13.csv"))%>%
  mutate(hydro = "2022MED")
wy <- read_csv(here(input_data_path, "wytype_may.csv")) %>%
  mutate(year = WY) %>% select(-WY)
outputa <- read_csv(here(output_path, "summarized_output/abundance_meanmin_all_alts_AdjHist.csv"))%>%
  mutate(hydro = "AdjHist")
outputb <- read_csv(here(output_path, "summarized_output/abundance_meanmin_all_alts_2022MED.csv"))%>%
  mutate(hydro = "2022MED")
```

::: {.panel-tabset}
## X2
```{r, include = FALSE}
# Combine data
x2 <- bind_rows(x2a, x2b)

x2_long <- x2 %>%
  pivot_longer(cols = 3:14, values_to = "x2", names_to = "month") %>%
  mutate(month = as.numeric(month)) %>%
  mutate(month_abb = factor(month.abb[month], levels = month.abb)) %>%
  mutate(scenario = forcats::fct_relevel(scenario,  c("StatusQuo", "MaxDS_Even","MaxDS_Hist", "SummerFall_Even", "Summer_Even", "Summer_Even_AltSMSCG",
                                                      "SummerFall_Hist", "Summer_Hist", "June", "MaxWater", "MaxWater_noSMSCG"))) 

x2_summerfall <- x2_long %>% filter(month > 5 & month<11)
```

### X2 by year, alternative, and hydrology
```{r, include = FALSE}
#| fig-width: 12
#| fig-height: 8
plot_x2 <- ggplot() +
    # geom_point(data = x2_summerfall, aes(x = year, y = x2, color = scenario, shape = scenario))+
    geom_line(data = x2_summerfall, aes(x = year, y = x2, color = scenario, linetype = scenario), alpha = 0.9)+
    # scale_shape_manual(values = c(20, 0, 18, 6, 8, 9, 10, 23, 1, 13, 14))+
    viridis::scale_color_viridis(discrete = TRUE, option = "turbo") + 
    scale_x_continuous(breaks = seq(1995, 2014, 1))+ 
    scale_y_continuous(breaks = seq(55, 95, 10)) + 
    facet_grid(month_abb~hydro) + 
    labs(y = "X2 (km)")+
    theme_bw() +
    theme(axis.text  = element_text(size = 11),
          axis.text.x = element_text(angle = 90))
```

```{r , echo = FALSE}
ggplotly(plot_x2)
```

## Zooplankton
```{r , include = FALSE}
### Combine data 
zoop <- bind_rows(zoopa, zoopb)
zoop_long <- zoop %>%
  pivot_longer(cols = starts_with("sal"),
               names_to = "alt",
               values_to = "zoop") %>%
  mutate(alt = str_remove(alt, "_median"),
         alt = str_remove(alt, "sal_")) %>%
  left_join(wy)
```

```{r, include = FALSE}
### Summarize by year
zoop_summ <- zoop_long %>%
  group_by(year, IBMR, hydro, alt, wytype) %>% 
  summarize(mnzoop = mean(zoop, na.rm = TRUE),
            sumzoop = sum(zoop, na.rm = TRUE)) %>%
  mutate(actionyr = if_else(wytype %in% c("AN", "W"), "*", "")) %>%
  filter(alt !="base") %>%
  rename(taxa = IBMR)

# Summarize by month
zoop_summ_month <- zoop_long %>%
  mutate(actionyr = if_else(wytype %in% c("AN", "W"), "*", "")) %>%
  group_by(month, IBMR, hydro, alt, actionyr) %>% 
  summarize(mnzoop = mean(zoop, na.rm = TRUE),
            sumzoop = sum(zoop, na.rm = TRUE)) %>%
  
  filter(alt !="base") %>%
  rename(taxa = IBMR)

zoop_summ_month_b <- zoop_long %>%
  mutate(actionyr = if_else(wytype %in% c("AN", "W"), "*", "")) %>%
  filter(year != 1997) %>%
  group_by(month, IBMR, hydro, alt, actionyr) %>% 
  summarize(mnzoop = mean(zoop, na.rm = TRUE),
            sumzoop = sum(zoop, na.rm = TRUE)) %>%
  
  filter(alt !="base") %>%
  rename(taxa = IBMR)
```

### Zooplankton by year (all years and alts)

*Adjusted Historical Hydrology*
*Action years marked by stars*

```{r, echo = FALSE}
#| fig-width: 12
#| fig-height: 8
ggplotly(ggplot(zoop_summ %>%filter(hydro == "AdjHist")) +
  geom_col(aes(x = factor(year), y = sumzoop, fill = taxa)) +
  geom_text(aes(x = factor(year), y = sumzoop + 5, label = actionyr))+
  labs(title = "AdjHist")+
  viridis::scale_fill_viridis(option = "turbo", discrete = TRUE)+
  facet_wrap(~alt)+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)))
```

*2022MED Hydrology*

```{r, echo = FALSE}
#| fig-width: 12
#| fig-height: 8
ggplotly(ggplot(zoop_summ %>%filter(hydro == "2022MED")) +
  geom_col(aes(x = factor(year), y = sumzoop, fill = taxa)) +
  geom_text(aes(x = factor(year), y = sumzoop + 5, label = actionyr))+
  facet_wrap(~alt)+
  labs(title = "2022MED")+
  viridis::scale_fill_viridis(option = "turbo", discrete = TRUE)+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)))
```

### Zooplankton by year (no 1997; all alts)

*Adjusted Historical Hydrology*
*Action years marked by stars*

```{r, echo = FALSE}
#| fig-width: 12
#| fig-height: 8
ggplotly(ggplot(zoop_summ %>%filter(hydro == "AdjHist", year != 1997)) +
   geom_col(aes(x = factor(year), y = sumzoop, fill = taxa)) +
   geom_text(aes(x = factor(year), y = sumzoop + 5, label = actionyr))+
   labs(title = "AdjHist no 1997")+
   viridis::scale_fill_viridis(option = "turbo", discrete = TRUE)+
   facet_wrap(~alt)+
   theme_bw() +
   theme(axis.text.x = element_text(angle = 90)))
```

*2022MED Hydrology*

```{r, echo = FALSE}
#| fig-width: 12
#| fig-height: 8
ggplotly(ggplot(zoop_summ %>%filter(hydro == "2022MED", year !=1997)) +
    geom_col(aes(x = factor(year), y = sumzoop, fill = taxa)) +
    geom_text(aes(x = factor(year), y = sumzoop + 5, label = actionyr))+
    facet_wrap(~alt)+
    labs(title = "2022MED no 1997")+
    viridis::scale_fill_viridis(option = "turbo", discrete = TRUE)+
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90)))
```

### Zooplankton by month


### Zooplankton by month, all years
*Filtered to action years (wet and an)*

```{r}
#| fig-height: 12
#| fig-width: 8
ggplotly(ggplot(zoop_summ_month %>%filter(actionyr == "*")) +
    geom_point(aes(x = factor(month), y = mnzoop, color = alt, shape = alt), size = 0.8) +
    facet_grid(taxa~hydro)+
    viridis::scale_color_viridis(option = "turbo", discrete = TRUE)+
    scale_shape_manual(values = c(20, 0, 18, 6, 8, 9, 10, 23, 1, 13, 14))+
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90)))
```
### Zooplankton by month, no 1997
*Filtered to action years (wet and an)*

```{r}
#| fig-height: 12
#| fig-width: 8
ggplotly(ggplot(zoop_summ_month_b %>%filter(actionyr == "*")) +
    geom_point(aes(x = factor(month), y = mnzoop, color = alt, shape = alt), size = 0.8) +
    facet_grid(taxa~hydro)+
    viridis::scale_color_viridis(option = "turbo", discrete = TRUE)+
    scale_shape_manual(values = c(20, 0, 18, 6, 8, 9, 10, 23, 1, 13, 14))+
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90)))
```




## Final Outputs

```{r, include = FALSE}
# Combine two hydros
output <- bind_rows(outputa, outputb)
```

### Lambda across alternatives and hydrologies
```{r, echo = FALSE}
#| fig-width: 7
#| fig-height: 6
ggplotly(ggplot(output) + 
    geom_point(aes(x =Alternatives, y = lambda, shape = hydro, color = hydro), size = 3)+
    geom_hline(yintercept = 1, linetype = "dashed") + 
    # geom_label(aes(x = Alternatives, y = MeanMinAbundance + 200000, label = round(MeanMinAbundance)), size = 3.5) + 
      scale_color_manual(values = c("navy", "steelblue4"))+
    labs(y = "Lambda")+ 
    theme_bw()+
    theme(axis.text.x = element_text(angle = 90, hjust = 0.99),
          axis.title = element_text(size = 12),
          axis.text = element_text(size = 12),
          axis.title.x = element_blank()))
```

### Minimum Abundance across alternatives and hydrologies
```{r, echo = FALSE}
#| fig-width: 7
#| fig-height: 6
ggplotly(ggplot(output) + 
    geom_point(aes(x =Alternatives, y = MeanMinAbundance, shape = hydro, color = hydro), size = 3)+
    geom_hline(yintercept = 500000, linetype = "dashed") + 
    # geom_label(aes(x = Alternatives, y = MeanMinAbundance + 200000, label = round(MeanMinAbundance)), size = 3.5) + 
      scale_color_manual(values = c("navy", "steelblue4"))+
    labs(y = "Mean Minimum Abundance")+ 
    theme_bw()+
    theme(axis.text.x = element_text(angle = 90, hjust = 0.99),
          axis.title = element_text(size = 12),
          axis.text = element_text(size = 12),
          axis.title.x = element_blank()))
```

:::